Select-based random access to variable-byte encodings

### 2020-03-04 gradu/msc_english_contents.tex
Optimized 4bit more

If the original element size is 32 bits and the data was compressed to k blocks of length b. The starting byte s of the element is $select$(i) * b div 8 where div is the integer division. 
Offset o is the remainder of previous division, o = x*b mod 8. The 32-bit word is read from memory from byte location s. Then bit shift left for 32-b*k-o is applied to remove trailing 
bits and bit shift right 32-b*k to re-align bits. If block size is one byte, offset calculation is not needed and thus the process is slightly faster.

@@

 - subarray access(?)

>>>>

For example, let the integer bit length be 32 and the original data be compressed to 5 blocks of 4 bits. First, the byte location of the first block is needed. $Select_1$(i) gets the index location
of needed block. Because the block length differs from byte length, the the two values need to be calculated. First, the starting byte location $s$ needs to be calculated 
$(s = select_1(i) \times 4) \div 8$, the latter operand being the integer division. Second, the first block might not be at the start of the byte $s$. This is why offset $o = select_1(i) \times 4) mod 8$
is needed (in the example case, o equals to either 0 or 4). Then a 32-bit word $w$ is read from memory from byte location $s$.

At this point, $w$ contains the wanted bits, but also has extra bits from the previous and next compressed integers. If offset was not zero, there are trailing $o$ bits in $w$ from the previous integer. 
These can be conveniently removed by bit-shifting $w$ right for $o$ bits. Then a pre-calculated bitmask (0xFFFFF in this case) is applied and $w$ contains the requested value. Because most systems are
able to read memory only from byte addresses, there is a corner case where the value is stored in too many bytes for a word. In this setup, this happens if compressed length was 8 blocks and offset was 4.
Then another step of reading another byte and storing the remaining bits to $w$ is required. This is entirely avoidable if integer length is constrained. In this case maximum integer length would be 28 bits.

@@

 - See Ahn \& Moffat 2005; Lemire et al. 2018
 - subarray access(?)

 - compare to elias-fano?
